|

### 2019-09-06 analysis/paper/manuscript.Rmd
Modifications for initial submission to Fish and Fisheries
Integrate Bob's paragraph and figures on his RStrap dashboard

---
title: |
  Improving the communication and accessibility of stock assessment using interactive visualization tools
author: "Paul M. Regular, Gregory J. Robertson, Keith P. Lewis"
date: |
  Fisheries and Oceans Canada, Northwest Atlantic Fisheries Center, 80 East White Hills, St. John's, Newfoundland and Labrador, A1C 5X1, Canada
output: 
  bookdown::word_document2:
    reference_docx: template.docx
bibliography: references.bib
csl: apa.csl
abstract: |
  **Running title**: Interactive stock assessment

@@

@@

@@

The data sets used in stock assessments are constantly growing due to the continuation of long-term monitoring efforts, the addition of new monitoring programs, or both. As such, stock assessment biologists need to manage large volumes of data from a variety of sources. Time series of reported landings and catch-at-age are often analyzed in conjunction with data from “fishery-independent” surveys that track changes in abundance and, in many cases, also monitor trends in biological factors such as age composition, growth rates, sex ratios and maturation stages. For some data-rich stocks, mark and recapture studies are also carried out to estimate movement, migration, growth rate, natural mortality, and discard mortality. All of the above-mentioned data sets are complex and as the volume and variety of these data increases, it becomes more difficult to be aware of the details of each data set and discover key patterns within each. This challenge can be mediated, to a degree, by the application of interactive visualization tools as that they allow detailed exploration of the data behind a plot. For instance, the ability to zoom in on features or areas of interest, turn off layers and hover over specific points to reveal more information creates an interactive user-driven experience that expedites explorations of data. This is exemplified by an interactive mapping tool developed for the exploration of a long-term tagging study.

@@

```{r map, echo = FALSE, fig.cap = '- Screen shot of cod tag mapping tool using shinydashboard and leaflet. The markercluster function dynamically splits or pools tagging locations (red, orange, green or yellow points) depending on zoom level, the recoveries positions (blue) are much fewer, and are left to be plotted individually at all scales. Options to include pop up labels are included, so specific information on each point can be retrieved with a mouse click (in this case: tag number, fish length, date released, and date captured), which is particularly useful when error checking.'}
knitr::include_graphics("figures/tagging_dashboard_map_tab.png")
```

```{r summary, echo = FALSE, fig.cap = '- Basic summaries of the recovery data from the tags selected within the shiny dashboard cod tag mapping tool. In this case, histograms and summary statistics of the recovery positions are returned, along with a simple map of kernels showing the 2D spatial distribution of the selected tag recoveries.'}
knitr::include_graphics("figures/tagging_dashboard_summary_tab.png")
```

@@

In the pursuit of an easier and more efficient way to communicate results from NCAM, an interactive and self-contained “dashboard” was developed for the 2018 assessment of Northern cod [@dwyer2019]. The concept of using a dashboard was borrowed from the business community where dashboards are frequently used to group a series of interactive visuals and tables to provide at-a-glance views of key performance indicators. We used R-based packages [@R] and the RStudio IDE [@Rstudio] to construct a tool for exploring the input and output of NCAM, specifically the flexdashboard [@allaire2017] package to group interactive plotly-based [@sievert2018] visuals into a dynamic document. We also used the crosstalk [@cheng2016] package to link the data displayed across multiple plots. Via R Markdown [@allaire2018], the dashboard is rendered into a self-contained html file that is reproducible, interactive, and easy to update following modifications to the model or the addition of new data.

@@

```{r rv, echo = FALSE, fig.cap = '- Screenshot of the "RV survey" tab from the NCAM dashboard where total observed (dots) and NCAM model predicted values (lines) for the DFO RV survey index are shown in the upper left panel and scaled matrix plot of age-disaggregated standardized log residuals are shown in the lower left panel (blue = positive, red = negative, symbols scaled by size; grey = index values of zero). Age-disaggregated observed (dots) and predicted (lines) values from the DFO RV survey are shown in the right panel.'}
knitr::include_graphics("figures/NCAM_dashboard_rv-survey_tab.png")
```

```{r trends, echo = FALSE, fig.cap = '- Screenshot of the "Trends" tab which displays estimates of recruitment, stock size, and stock size relative to B~lim~ (left panels) and mortality rates (F, M, Z, right panels).'}
knitr::include_graphics("figures/NCAM_dashboard_trends_tab.png")
```

@@

The amount of data available to scientists has grown by orders of magnitude in recent decades as has the complexities of data management, exploratory data analysis, formal analyses and associated diagnostics [@lewis2018]. The majority of this sequence of events, sometimes called “the data pipeline” [@leek2015a], have not traditionally been part of the peer-review process which sees only the end products of an analysis. However, decisions made along the data pipeline increasingly influence the outcome of the study. Gelman and Loken [-@gelman2014] coined the term “garden of forking paths” to illustrate that different conclusions can be arrived at depending on what decisions are made along during different stages of the analysis. Due to a number of limitations, such as available pages in journals, much of the data pipeline is not transparent nor is it reproducible. A number of authors have recently advocated for a culture of open science and reproducible research, i.e., a change in the transparency and reproducibility of science [@leek2015a; @leek2015b; @hampton2013; @hampton2015]. Proponents of open science and reproducible research highlight a number of benefits including a more productive and responsible scientific culture, an ability to address larger and more complex questions, as well as a more efficient workflow and ability to reproduce one’s own work [@fomel2009].

>>>>

---
title: |
  Improving the communication and accessibility of stock assessment using interactive visualization tools
author: "Paul M. Regular, Gregory J. Robertson, Robert Rogers, Keith P. Lewis"
date: |
  Fisheries and Oceans Canada, Northwest Atlantic Fisheries Center, 80 East White Hills, St. John's, Newfoundland and Labrador, A1C 5X1, Canada
output: 
  bookdown::word_document2:
    reference_docx: template.docx
bibliography: references.bib
csl: apa.csl
abstract: |
  **Running title**: Interactive stock assessment

@@

  **Data Availability Statement**: Data sharing is not applicable to this article as no new data were created or analyzed in this study.

@@

<!-- READ AND REFERENCE: -->
<!-- - https://www.sciencedirect.com/science/article/pii/S1364815210000551 -->

<!-- AIM FOR SHORT COMMUNICATION IN Environmental Modeling & Software -->

@@

The data sets used in stock assessments are constantly growing due to the continuation of long-term monitoring efforts, the addition of new monitoring programs, or both. As such, stock assessment biologists need to manage large volumes of data from a variety of sources. Time series of reported landings and catch-at-age are often analyzed in conjunction with data from "fishery-independent" surveys that track changes in abundance and, in many cases, also monitor trends in biological factors such as age composition, growth rates, sex ratios and maturation stages. For some data-rich stocks, mark and recapture studies are also carried out to estimate movement, migration, growth rate, natural mortality, and discard mortality. All of the above-mentioned data sets are complex and as the volume and variety of these data increases, it becomes more difficult to be aware of the details of each data set and discover key patterns within each. This challenge can be mediated, to a degree, by the application of interactive visualization tools as that they allow detailed exploration of the data behind a plot. For instance, the ability to zoom in on features or areas of interest, turn off layers and hover over specific points to reveal more information creates an interactive user-driven experience that expedites explorations of data. This is exemplified by two recently developed interactive tools: 1) a tool designed to quickly examine "fishery independent" survey data; and 2) an interactive  mapping tool developed for the exploration of a long-term tagging study.

<!-- should the concept of a dashboard be introduced here? -->

Fisheries and Oceans Canada (DFO) has been conducting a multi-species stratified-random survey across the Newfoundland and Labrador shelf since the 1970s [@rideout2018]. Currently, these data are analyzed using a standard stratified analysis via a locally developed R package called RStrap [for details on methodology see @smith1981]. Both the inputs and outputs from the analysis are very large and, depending on the species, the time-series may include > 40 years and span the majority of the Newfoundland and Labrador shelf. In order for people to quickly and reliability explore these data without iteratively modifying R scripts, an app was built using a combination of shiny [@chang2018a], crosstalk [@cheng2016], flexdashboard [@allaire2017], plotly [@sievert2018], and R Markdown [@allaire2018] all within the framework of the R programming language [@R] and the RStudio IDE [@Rstudio]. The app started as a way to visualize estimates of biomass and abundance trends of specific species using the flexdashboard package, however, shiny was latter applied as it allows stratified analyses to be run in the background and therefore allows the user to dynamically explore estimates from multiple species in the one session. Both crosstalk and plotly were incorporated to allow the user to interact with the data. The app contains four primary tabs ("Survey Indices", "Age & Length Distributions", "Recruitment", and "Help") containing stock level estimates of biomass (Figure \@ref(fig:biomass)) and abundance, mean numbers and weights per standardized survey tow, stock level biomass broken out by division, age and length frequencies (Figure \@ref(fig:freq)), and recruitment indices. The number of potential inputs to RStrap are rather large and most do not need to be modified. Some inputs however, do require modification. The input interface on the RStrap Explorer app is divided into two parts: basic and advanced inputs. Basic inputs are those like species, season, and survey year(s). Advanced inputs allow the user to select unconverted data (default) or converted data (which allows for a single complete time-series to be used as opposed to time-series broken by changes in survey gear) and select whether they would like analyses conducted by length and/or age, among others. This app was built for data exploration and allow the user to obtain quick and easy visualization of survey data, but is not meant to replace the through data checking and analysis required for formal stock assessment.

```{r biomass, echo = FALSE, fig.cap = '- Screen shot of RStrap Explorer tool, using flexdashboard, plotly, and shiny. On the left side of the figure, one can see the "Basic Inputs" sidebar, which is fixed across all pages and allows the user to specify the species, NAFO Divisions, season, and survey years of interest. The "Biomass Index" sub-tab is illustrated here. This sub-tab shows the unconverted time-series of biomass data with 95% confidence intervals (Engle and Campelen) for both cod (*Gadus morhua*; species code 438) and American Plaice (*Hippoglossoides platessoides*; species code 889). All trend lines can be toggled off using plotly’s dynamic user interface. The user also has the ability to hover over the points to compare estimates and export the figure as a png.'}
knitr::include_graphics("figures/RStrap_dashboard_biomass.png")
```

```{r freq, echo = FALSE, fig.cap = '- "Age and Length Frequencies" tab of the RStrap Explorer tool. This tab illustrates the length frequency distributions of both cod (*Gadus morhua*; species code 438) and American plaice (*Hippoglossoides platessoides*; species code 889) binned into 2 cm length bins. The binning is easily modified using the "Length Group" field in the "Advanced Options" drop down menu (shown on left side).'}
knitr::include_graphics("figures/RStrap_dashboard_freq.png")
```

@@

```{r map, echo = FALSE, fig.cap = 'Screen shot of cod tag mapping tool using shinydashboard and leaflet. The markercluster function dynamically splits or pools tagging locations (red, orange, green or yellow points) depending on zoom level, the recoveries positions (blue) are much fewer, and are left to be plotted individually at all scales. Options to include pop up labels are included, so specific information on each point can be retrieved with a mouse click (in this case: tag number, fish length, date released, and date captured), which is particularly useful when error checking.'}
knitr::include_graphics("figures/tagging_dashboard_map_tab.png")
```

```{r summary, echo = FALSE, fig.cap = 'Basic summaries of the recovery data from the tags selected within the shiny dashboard cod tag mapping tool. In this case, histograms and summary statistics of the recovery positions are returned, along with a simple map of kernels showing the 2D spatial distribution of the selected tag recoveries.'}
knitr::include_graphics("figures/tagging_dashboard_summary_tab.png")
```

@@

In the pursuit of an easier and more efficient way to communicate results from NCAM, an interactive and self-contained "dashboard" was developed for the 2018 assessment of Northern cod [@dwyer2019]. The concept of using a dashboard was borrowed from the business community where dashboards are frequently used to group a series of interactive visuals and tables to provide at-a-glance views of key performance indicators. We used R-based packages [@R] and the RStudio IDE [@Rstudio] to construct a tool for exploring the input and output of NCAM, specifically the flexdashboard [@allaire2017] package to group interactive plotly-based [@sievert2018] visuals into a dynamic document. We also used the crosstalk [@cheng2016] package to link the data displayed across multiple plots. Via R Markdown [@allaire2018], the dashboard is rendered into a self-contained html file that is reproducible, interactive, and easy to update following modifications to the model or the addition of new data.

@@

```{r rv, echo = FALSE, fig.cap = 'Screenshot of the "RV survey" tab from the NCAM dashboard where total observed (dots) and NCAM model predicted values (lines) for the DFO RV survey index are shown in the upper left panel and scaled matrix plot of age-disaggregated standardized log residuals are shown in the lower left panel (blue = positive, red = negative, symbols scaled by size; grey = index values of zero). Age-disaggregated observed (dots) and predicted (lines) values from the DFO RV survey are shown in the right panel.'}
knitr::include_graphics("figures/NCAM_dashboard_rv-survey_tab.png")
```

```{r trends, echo = FALSE, fig.cap = 'Screenshot of the "Trends" tab which displays estimates of recruitment, stock size, and stock size relative to B~lim~ (left panels) and mortality rates (F, M, Z, right panels).'}
knitr::include_graphics("figures/NCAM_dashboard_trends_tab.png")
```

@@

The amount of data available to scientists has grown by orders of magnitude in recent decades as has the complexities of data management, exploratory data analysis, formal analyses and associated diagnostics [@lewis2018]. The majority of this sequence of events, sometimes called "the data pipeline" [@leek2015a], have not traditionally been part of the peer-review process which sees only the end products of an analysis. However, decisions made along the data pipeline increasingly influence the outcome of the study. Gelman and Loken [-@gelman2014] coined the term "garden of forking paths" to illustrate that different conclusions can be arrived at depending on what decisions are made along during different stages of the analysis. Due to a number of limitations, such as available pages in journals, much of the data pipeline is not transparent nor is it reproducible. A number of authors have recently advocated for a culture of open science and reproducible research, i.e., a change in the transparency and reproducibility of science [@leek2015a; @leek2015b; @hampton2013; @hampton2015]. Proponents of open science and reproducible research highlight a number of benefits including a more productive and responsible scientific culture, an ability to address larger and more complex questions, as well as a more efficient workflow and ability to reproduce one’s own work [@fomel2009].
